Sulfur Dioxide on Water's Surface

### 2011-03-18 Background.tex
Writing about the density distro/profile, and the way to claculate the surface layer

	The results of analyses performed in this study are related to the location of the liquid water surface. A surface, however, is a dynamic boundary that is neither flat or stationary. Several previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used to extract fitting parameters, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface will change, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformation the surface undergoes. In this work we approach the problem of surface loation by calculating a reference location for each timestep by averaging the long-axis positions of the waters contained in the topmost monolayer. This provides us with a consistent reference point in the simulations to which we relate our analyses. This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.

  The location of the water surface was thus determined by the averaged positions of a set number of the topmost waters in the simulated aqueous slabs. The number of waters counted was decided by taking a few issues into account. First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer. Alternatively, assuming a spherical model of water with a radius of 2.4\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a maximum of 52 molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness. Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.

@@

\subsection{Hydrate Complex Analysis}

	During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule. Computationally, there are various methods to determine hydrate complex formation. The most simple metric is to count the number of hydrogen bonds formed from the \suldiox~to neighboring molecules. This method can be furthered extended to also include the strength of those bonding interactions by additionally reporting the bondlengths to the hydrogen bonding partners. Longer bondlengths are typical of weaker bonds, and this method is used in this work to identify the bonding environment of the \suldiox~during solvation.

	A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox. However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form. In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries. Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works. As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.

	We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems. The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph. Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex. We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph. After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black. During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph. At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
		\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges. The bonds are all given equal weights and the graph is undirected. The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered. A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4). At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle. This cycle is composed of 8 atoms, and 3 molecules}
		\label{fig:so2-complex-graph}
	\end{center}
\end{figure}

>>>>

	The results of analyses performed in this study are related to the location of the liquid water surface to relate depth profiles of the orientation distributions. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we approach the problem of defining surface location by calculating a reference location for each timestep. We perform an averaging of the positions of the waters contained in the topmost monolayer. This provides us with a consistent reference point in the simulations to which we relate our analyses. This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.

  The location of the water surface was determined by the averaged positions of a set number of the outter-most waters in the simulated aqueous slabs. The number of waters included in the averaging was decided by taking a few issues into account. First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer. Alternatively, assuming a spherical model of water with a radius of 2.2\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work. The water density profile (dashed line) is produced by averaging the density profile of each timestep in the simulation. The water profile was then fit with a $\tanh$ function lineshape to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$. A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the surface waters, or the entire slab over the course of the trajectory. Thus the calculated location and width using the fitting routine are not accurate over long trajectories.

The new procedure described above for finding the outter monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue). It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.

@@

%\subsection{Hydrate Complex Analysis}

%During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule. Computationally, there are various methods to determine hydrate complex formation. The most simple metric is to count the number of hydrogen bonds formed from the \suldiox~to neighboring molecules. This method can be furthered extended to also include the strength of those bonding interactions by additionally reporting the bondlengths to the hydrogen bonding partners. Longer bondlengths are typical of weaker bonds, and this method is used in this work to identify the bonding environment of the \suldiox~during solvation.

%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox. However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form. In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries. Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works. As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.

%We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems. The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph. Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex. We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph. After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black. During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph. At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.

%\begin{figure}[h!]
%\begin{center}
%\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
%\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges. The bonds are all given equal weights and the graph is undirected. The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered. A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4). At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle. This cycle is composed of 8 atoms, and 3 molecules}
%\label{fig:so2-complex-graph}
%\end{center}
%\end{figure}
